---
title: "Hotspot enrichment analysis"
author: Ahmed Hasan
output: pdf_document
---

Questions:
1. Which annotations in the genome are enriched for hotspots?

```{r setup, include=FALSE}
require(knitr)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r}
library(tidyverse)
library(fs)
library(broom)
```

## Loading in data

```{r}
fnames <- dir_ls('data/correlates/hotspot_enrichment/', regexp = 'chromosome_[0-9]+\\.txt')
annot <- map_dfr(fnames, read_delim, delim = ' ', col_types = cols(), .id = 'chrom') %>% 
  mutate(chrom = str_extract(chrom, 'chromosome_[0-9]+'))

annot_genome <- annot %>% 
  select(-chrom) %>% 
  group_by(correlate) %>% 
  summarise_all(sum)

```

## Enrichment

Let's start with the 'upstream' annotation. We're asking whether upstream is enriched in our hotspot regions, compared to the background.

$x$ = count of upstream sites in hotspots


| type | upstream | non-upstream |
| ----- |-----|-----|
| hotspot | x | total_hotspots_all - x |
| non-hotspot | total_upstream - x | total_non_upstream - total_hotspots_all + x |
| total | total_upstream | total_non_upstream |

or

| type | upstream | non-upstream |
| ----- |-----|-----|
| hotspot | x | total_hotspots_all - x |
| non-hotspot | total_upstream - x | total_non_hotspot_non_upstream + x |
| total | total_upstream | total_non_upstream |


```{r}
calculate_enrichment <- function(d, input_correlate) {
  x <- d[d$correlate == input_correlate, ]$hotspot_count
  total_correlate <- d[d$correlate == input_correlate, ]$total_count
  total_hotspots_all <- sum(d$hotspot_count)
  total_non_correlate <- d %>% 
    filter(correlate != input_correlate) %>% 
    select(total_count) %>% 
    sum()
  total_hotspot_non_correlate <- d %>% 
    filter(correlate != input_correlate) %>% 
    select(hotspot_count) %>% 
    sum()
  total_non_hotspot_non_correlate <- total_non_correlate - total_hotspot_non_correlate
  
  fisher_matrix <- matrix(
    c(x, total_correlate - x,
      total_hotspots_all - x, total_non_hotspot_non_correlate + x),
    2, 2
  )
  
  t <- fisher.test(fisher_matrix, alternative = 'greater')
  return(t)
  
}

```


Over all correlates:


```{r}
correlates <- as.list(annot_genome$correlate)
names(correlates) <- annot_genome$correlate

correlates %>% 
  map_dfr(~ 
    calculate_enrichment(annot_genome, .) %>% 
    broom::tidy(), .id = 'correlate'
  )
```

## Percent of hotspots contained in annotations

```{r}
annot_genome %>% 
  mutate(prop_hotspots = hotspot_count / sum(hotspot_count),
         prop_total = total_count / sum(total_count)) %>% 
  filter(correlate %in% c('upstream')) %>% 
  summarise_at(vars(contains('prop')), sum)
```



