---
title: "Intergenic tract analysis"
author: Ahmed Hasan
output: pdf_document
---

```{r}
library(tidyverse)
library(fs)
library(here)


din1 <- 6.34375
din2 <- 3.385417
```

## Load data

```{r}
d <- here('data/correlates/intergenic_tract_rho.tsv') %>% 
  read_tsv(col_types = cols()) %>% 
  filter(tract_size > 0)
```

Schematic:
- less than 2 kb ('short')
  - both = L
- 2 kb - 4 kb ('medium')
  - D = L - 2
  - upstream = downstream = D
  - both = L - 2D
- greater than 4 kb ('long')
  - upstream = downstream = 2 kb
  - intergenic = L - 4
  - both = 0
- greater than 8 kb ('extended')
  - upstream = downstream = 2 kb
  - intergenic = L - 4
  - both = 0
  - intergenic > upstream/downstream
  
  
```{r}
d_bins <- d %>% 
  mutate(
    bin = case_when(
      tract_size <= 2000 ~ 'short',
      tract_size > 2000 & tract_size <= 4000 ~ 'medium',
      tract_size > 4000 & tract_size <= 8000 ~ 'long',
      tract_size > 8000 ~ 'extended'
    )
  ) %>% 
  mutate(
    both_sites = case_when(
      bin == 'short' ~ tract_size,
      bin == 'medium' ~ tract_size - (2 * (tract_size - 2000)),
      bin == 'long' ~ 0,
      bin == 'extended' ~ 0,
    ),
    upstream_sites = case_when(
      bin == 'short' ~ 0,
      bin == 'medium' ~ tract_size - 2000,
      bin == 'long' ~ 2000,
      bin == 'extended' ~ 2000,
    ),
    intergenic_sites = case_when(
      bin == 'short' ~ 0,
      bin == 'medium' ~ 0,
      bin == 'long' ~ tract_size - 4000,
      bin == 'extended' ~ tract_size - 4000
    )
  ) %>% 
  mutate(downstream_sites = upstream_sites)
```

## Distributions

Overall tract sizes:

```{r}
ggplot(d_bins, aes(x = log10(tract_size))) +
  geom_histogram() +
  theme_bw()

ggplot(d_bins, aes(x = log10(tract_size))) +
  stat_ecdf() +
  theme_bw()
```


Tract sizes by annotation:

```{r}
d_bins %>% 
  select(ends_with('sites')) %>% 
  summary()
```

Looks how we'd expect, with both/upstream/downstream capped at 2000 and intergenic being much larger

Tract counts by bin:

```{r}
d_bins %>% 
  group_by(bin) %>% 
  tally() %>% 
  mutate(prop = n / sum(n))
```

Short tracts make up *90.9 percent* of all intergenic tracts! 

Rho distributions by tract size:

```{r}
d_bins %>% 
  ggplot(aes(x = bin, y = log10(rho_window))) +
  geom_boxplot() +
  geom_jitter(alpha = 0.2)
```

```{r}
d_bins %>% 
  group_by(bin) %>% 
  summarise(
    total_rho = sum(rho_vals),
    total_count = sum(rho_count),
    sd_rho = sd(rho_window),
    n = n()
  ) %>% 
  ungroup() %>% 
  mutate(
    mean_rho = total_rho / total_count,
    se_rho = sd_rho / sqrt(n)
  )
```

Visualizing this:

```{r}
d_summarised <- d_bins %>% 
  group_by(bin) %>% 
  summarise(
    total_rho = sum(rho_vals),
    total_count = sum(rho_count),
    sd_rho = sd(rho_window),
    n = n()
  ) %>% 
  ungroup() %>% 
  mutate(
    mean_rho = total_rho / total_count,
    se_rho = sd_rho / sqrt(n)
  ) %>% 
  mutate(prop = n / sum(n)) %>% 
  select(bin, prop, n, mean_rho, se_rho)

d_summarised %>% 
  ggplot(aes(x = bin, y = mean_rho)) +
  geom_bar(stat = 'identity') +
  geom_errorbar(
    aes(ymax = mean_rho + se_rho, ymin = mean_rho - se_rho),
    size = 0.4, width = 0.3
  ) +
  theme_classic()

d_summarised %>% 
  ggplot(aes(x = bin, y = n)) +
  geom_bar(stat = 'identity') +
  theme_classic()
```

The overall trend looks pretty messy though:

```{r}
ggplot(d_bins, aes(x = log10(tract_size), y = log10(rho_window))) +
  geom_point(alpha = 0.4, aes(color = bin)) +
  geom_smooth(method = 'lm')

```

Bin tract size on log scale and correlate with rho:

```{r}
d_bins_log <- d_bins %>% 
  mutate(log_size = log10(tract_size)) %>% 
  mutate(log_bin = floor(log_size)) %>% 
  group_by(log_bin) %>% 
  summarise(
    total_rho = sum(rho_vals),
    total_count = sum(rho_count),
    sd_rho = sd(rho_window),
    n = n()
  ) %>% 
  ungroup() %>% 
  mutate(
    mean_rho = total_rho / total_count,
    se_rho = sd_rho / sqrt(n)
  ) %>% 
  mutate(bin_start = as.integer(10 ^ log_bin),
         bin_end = lead(bin_start)) %>% 
  mutate(bin_end = ifelse(is.na(bin_end), Inf, bin_end)) %>% 
  mutate(bin_name = paste0(bin_start, '-', bin_end))

d_bins_log

d_bins_log %>% 
  filter(bin_start != 100000) %>% 
  ggplot(aes(x = bin_name, y = mean_rho)) +
  geom_point() +
  geom_errorbar(aes(ymin = mean_rho - se_rho, ymax = mean_rho + se_rho))
```



## Rho in UTRs

```{r}
utr <- here('data/correlates/utr_tract_rho.tsv') %>% 
  read_tsv(col_types = cols()) %>% 
  filter(tract_size > 0)

utr3 <- here('data/correlates/utr3_tract_rho.tsv') %>% 
  read_tsv(col_types = cols())

utr_tracts <- left_join(utr, d_bins, 
  by = c('chrom', 'start', 'end', 'tract_size', 'rho_vals', 'rho_count', 'rho_window'))

utr3_full <- left_join(utr3, d_bins,
  by = c('chrom', 'start', 'end')) %>% 
  filter(!is.na(bin))
```

How does rho vary in UTRs by bin?

```{r}
# 5' UTRs
utr_full %>% 
  group_by(bin) %>% 
  summarise(
    total_rho = sum(utr_rho_vals),
    total_count = sum(utr_rho_count),
    sd_rho = sd(utr_rho_window),
    n = n()
  ) %>% 
  ungroup() %>% 
  mutate(
    mean_rho = total_rho / total_count,
    se_rho = sd_rho / sqrt(n)
  )

# 3' UTRs
utr3_full %>% 
  group_by(bin) %>% 
  summarise(
    total_rho = sum(utr3_rho_vals),
    total_count = sum(utr3_rho_count),
    sd_rho = sd(utr3_rho_window),
    n = n()
  ) %>% 
  ungroup() %>% 
  mutate(
    mean_rho = total_rho / total_count,
    se_rho = sd_rho / sqrt(n)
  )
  
```

In the 'extended' bin (ie no instances of 'both' sites) - mean rho at adjacent UTRs is well above the genomewide average!

Otherwise, rho is super low - and especially so in the few 'medium' tracts. 

```{r}
utr_bars <- utr_full %>% 
  group_by(bin) %>% 
  summarise(
    total_rho = sum(utr_rho_vals),
    total_count = sum(utr_rho_count),
    sd_rho = sd(utr_rho_window),
    n = n()
  ) %>% 
  ungroup() %>% 
  mutate(
    mean_rho = total_rho / total_count,
    se_rho = sd_rho / sqrt(n)
  )

utr3_bars <- utr3_full %>% 
  group_by(bin) %>% 
  summarise(
    total_rho = sum(utr3_rho_vals),
    total_count = sum(utr3_rho_count),
    sd_rho = sd(utr3_rho_window),
    n = n()
  ) %>% 
  ungroup() %>% 
  mutate(
    mean_rho = total_rho / total_count,
    se_rho = sd_rho / sqrt(n)
  )

utr_bars_all <- bind_rows(utr_bars, utr3_bars, .id = 'name') %>% 
  mutate(utr = case_when(
    name == 1 ~ 'five_prime',
    name == 2 ~ 'three_prime')
  )

ggplot(utr_bars_all, aes(x = bin, y = mean_rho, fill = utr)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  geom_errorbar(width = 0.4, position = position_dodge(width = 0.9),
    aes(ymin = mean_rho - se_rho, 
        ymax = mean_rho + se_rho)) +
  theme_minimal()


```


How does this look for percentile bins of tract length?

```{r}
utr_full %>% 
  select(tract_size, utr_rho_window) %>%
  arrange(tract_size) %>% 
  mutate(tract_percentile = percent_rank(tract_size)) %>% 
  mutate(percentile_bin = floor(tract_percentile / 0.05) * 0.05) %>% 
  ggplot(aes(x = tract_percentile, y = utr_rho_window)) +
  geom_point() +
  geom_smooth()

```


## Hotspots in intergenic tracts

```{r}
intergenic_hotspots <- here('data/correlates/intergenic_tract_hotspots.tsv') %>% 
  read_tsv(col_types = cols()) %>% 
  filter(tract_size > 0) %>% 
  left_join(d_bins, by = c('chrom', 'start', 'end', 'rho_vals', 'rho_count', 'tract_size', 'rho_window'))
```

Hotspot counts/proportions by bin:

```{r}
intergenic_hotspots %>% 
  group_by(bin) %>% 
  summarise(
    avg_n_hotspots = mean(n_hotspots),
    avg_sites_in_hotspot = sum(sites_in_hotspot) / sum(rho_count),
    avg_upstream = mean(upstream_sites),
    avg_downstream = mean(downstream_sites),
    avg_both = mean(both_sites),
    avg_intergenic = mean(intergenic_sites)
  )
```

Hotspot strength:

```{r}
intergenic_hotspots %>% 
  filter(n_hotspots > 0) %>% 
  mutate(rate_ratio = rate_sum / n_hotspots) %>% 
  filter(rate_ratio < 2000) %>% # one freak hotspot
  group_by(bin) %>% 
  summarise(mean(rate_ratio))
```


















