---
title: "Intergenic tract analysis"
author: Ahmed Hasan
output: pdf_document
---

```{r}
library(tidyverse)
library(fs)
library(here)


din1 <- 6.34375
din2 <- 3.385417
```

## Load data

```{r}
d <- here('data/correlates/intergenic_tract_rho.tsv') %>% 
  read_tsv(col_types = cols()) %>% 
  filter(tract_size > 0)
```

Schematic:
- less than 2 kb ('short')
  - both = L
- 2 kb - 4 kb ('medium')
  - D = L - 2
  - upstream = downstream = D
  - both = L - 2D
- greater than 4 kb ('long')
  - upstream = downstream = 2 kb
  - intergenic = L - 4
  - both = 0
  
  
```{r}
d_bins <- d %>% 
  mutate(
    bin = case_when(
      tract_size <= 2000 ~ 'short',
      tract_size > 2000 & tract_size <= 4000 ~ 'medium',
      tract_size > 4000 ~ 'long'
    )
  ) %>% 
  mutate(
    both_sites = case_when(
      bin == 'short' ~ tract_size,
      bin == 'medium' ~ tract_size - (2 * (tract_size - 2000)),
      bin == 'long' ~ 0
    ),
    upstream_sites = case_when(
      bin == 'short' ~ 0,
      bin == 'medium' ~ tract_size - 2000,
      bin == 'long' ~ 2000
    ),
    intergenic_sites = case_when(
      bin == 'short' ~ 0,
      bin == 'medium' ~ 0,
      bin == 'long' ~ tract_size - 4000
    )
  ) %>% 
  mutate(downstream_sites = upstream_sites)
```

## Distributions

Overall tract sizes:

```{r}
ggplot(d_bins, aes(x = log10(tract_size))) +
  geom_histogram() +
  theme_bw()
```


Tract sizes by annotation:

```{r}
d_bins %>% 
  select(ends_with('sites')) %>% 
  summary()
```

Looks how we'd expect, with both/upstream/downstream capped at 2000 and intergenic being much larger

Tract counts by bin:

```{r}
d_bins %>% 
  group_by(bin) %>% 
  tally() %>% 
  mutate(prop = n / sum(n))
```

Short tracts make up *90.9 percent* of all intergenic tracts! 

Rho distributions by tract size:

```{r}
d_bins %>% 
  ggplot(aes(x = bin, y = log10(rho_window))) +
  geom_boxplot() +
  geom_jitter(alpha = 0.2)
```

```{r}
d_bins %>% 
  group_by(bin) %>% 
  summarise(
    total_rho = sum(rho_vals),
    total_count = sum(rho_count),
    sd_rho = sd(rho_window),
    n = n()
  ) %>% 
  ungroup() %>% 
  mutate(
    mean_rho = total_rho / total_count,
    se_rho = sd_rho / sqrt(n)
  )
```

Visualizing this:

```{r}
d_summarised <- d_bins %>% 
  group_by(bin) %>% 
  summarise(
    total_rho = sum(rho_vals),
    total_count = sum(rho_count),
    sd_rho = sd(rho_window),
    n = n()
  ) %>% 
  ungroup() %>% 
  mutate(
    mean_rho = total_rho / total_count,
    se_rho = sd_rho / sqrt(n)
  ) %>% 
  mutate(prop = n / sum(n)) %>% 
  select(bin, prop, n, mean_rho, se_rho)

d_summarised %>% 
  ggplot(aes(x = bin, y = mean_rho)) +
  geom_bar(stat = 'identity') +
  geom_errorbar(
    aes(ymax = mean_rho + se_rho, ymin = mean_rho - se_rho),
    size = 0.4, width = 0.3
  ) +
  theme_classic()

d_summarised %>% 
  ggplot(aes(x = bin, y = n)) +
  geom_bar(stat = 'identity') +
  theme_classic()
```

The overall trend looks pretty messy though:

```{r}
ggplot(d_bins, aes(x = log10(tract_size), y = log10(rho_window))) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = 'lm')

```

## Rho in UTRs

```{r}
utr <- here('data/correlates/utr_tract_rho.tsv') %>% 
  read_tsv(col_types = cols()) %>% 
  filter(tract_size > 0)

utr_full <- left_join(
  utr, d_bins, 
  by = c('chrom', 'start', 'end', 'tract_size', 'rho_vals', 'rho_count', 'rho_window'))
```

How does rho vary in UTRs by bin?

```{r}
utr_full %>% 
  group_by(bin) %>% 
  summarise(
    total_rho = sum(utr_rho_vals),
    total_count = sum(utr_rho_count),
    sd_rho = sd(utr_rho_window),
    n = n()
  ) %>% 
  ungroup() %>% 
  mutate(
    mean_rho = total_rho / total_count,
    se_rho = sd_rho / sqrt(n)
  )
```

In the 'long' bin (ie no instances of 'both' sites) - mean rho at adjacent UTRs is above the genomewide average!

Otherwise, rho is super low - and especially so in the few 'medium' tracts. 



