---
title: "Intergenic tract analysis"
author: Ahmed Hasan
output: pdf_document
---

```{r}
library(tidyverse)
library(fs)
library(here)


din1 <- 6.34375
din2 <- 3.385417
```

## Load data

```{r}
d <- here('data/correlates/intergenic_tract_rho.tsv') %>% 
  read_tsv(col_types = cols()) %>% 
  filter(tract_size > 0)
```

Schematic:
- less than 2 kb ('short')
  - both = L
- 2 kb - 4 kb ('medium')
  - D = L - 2
  - upstream = downstream = D
  - both = L - 2D
- greater than 4 kb ('long')
  - upstream = downstream = 2 kb
  - intergenic = L - 4
  - both = 0
- greater than 8 kb ('extended')
  - upstream = downstream = 2 kb
  - intergenic = L - 4
  - both = 0
  - intergenic > upstream/downstream
  
  
```{r}
d_bins <- d %>% 
  mutate(
    bin = case_when(
      tract_size <= 2000 ~ 'short',
      tract_size > 2000 & tract_size <= 4000 ~ 'medium',
      tract_size > 4000 & tract_size <= 8000 ~ 'long',
      tract_size > 8000 ~ 'extended'
    )
  ) %>% 
  mutate(
    both_sites = case_when(
      bin == 'short' ~ tract_size,
      bin == 'medium' ~ tract_size - (2 * (tract_size - 2000)),
      bin == 'long' ~ 0,
      bin == 'extended' ~ 0,
    ),
    upstream_sites = case_when(
      bin == 'short' ~ 0,
      bin == 'medium' ~ tract_size - 2000,
      bin == 'long' ~ 2000,
      bin == 'extended' ~ 2000,
    ),
    intergenic_sites = case_when(
      bin == 'short' ~ 0,
      bin == 'medium' ~ 0,
      bin == 'long' ~ tract_size - 4000,
      bin == 'extended' ~ tract_size - 4000
    )
  ) %>% 
  mutate(downstream_sites = upstream_sites)
```

## Distributions

Overall tract sizes:

```{r}
d_bins %>% 
  filter(tract_size < 10000) %>% 
  ggplot(aes(x = tract_size)) +
  geom_histogram() +
  theme_bw()

d_bins %>% 
  filter(tract_size < 10000) %>% 
  ggplot(aes(x = log10(tract_size))) +
  geom_histogram() +
  theme_bw()

# get proportion of genome for each intergenic bins
# is most of the intergenic sequence in the larger tracts? 

ggplot(d_bins, aes(x = log10(tract_size))) +
  stat_ecdf() +
  theme_bw()
```

```{r}
# tract size at 50th percentile

d_bins %>% 
  mutate(perc_rank = percent_rank(tract_size)) %>% 
  arrange(perc_rank) %>% 
  select(tract_size, perc_rank) %>% 
  filter(perc_rank > 0.498, perc_rank < 0.503) # 142 bp
```


Tract sizes by annotation:

```{r}
d_bins %>% 
  select(ends_with('sites')) %>% 
  summary()
```

Looks how we'd expect, with both/upstream/downstream capped at 2000 and intergenic being much larger

Tract counts by bin:

```{r}
d_bins %>% 
  group_by(bin) %>% 
  tally() %>% 
  mutate(prop = n / sum(n))
```

Short tracts make up *90.9 percent* of all intergenic tracts! 

Rho distributions by tract size:

```{r}
d_bins %>% 
  ggplot(aes(x = bin, y = log10(rho_window))) +
  geom_boxplot() +
  geom_jitter(alpha = 0.2)
```

```{r}
d_bins %>% 
  group_by(bin) %>% 
  summarise(
    total_rho = sum(rho_vals),
    total_count = sum(rho_count),
    sd_rho = sd(rho_window),
    n = n()
  ) %>% 
  ungroup() %>% 
  mutate(
    mean_rho = total_rho / total_count,
    se_rho = sd_rho / sqrt(n)
  )
```

Visualizing this:

```{r}
d_summarised <- d_bins %>% 
  group_by(bin) %>% 
  summarise(
    total_rho = sum(rho_vals),
    total_count = sum(rho_count),
    sd_rho = sd(rho_window),
    n = n()
  ) %>% 
  ungroup() %>% 
  mutate(
    mean_rho = total_rho / total_count,
    se_rho = sd_rho / sqrt(n)
  ) %>% 
  mutate(prop = n / sum(n)) %>% 
  select(bin, prop, n, mean_rho, se_rho)

d_summarised %>% 
  ggplot(aes(x = bin, y = mean_rho)) +
  geom_bar(stat = 'identity') +
  geom_errorbar(
    aes(ymax = mean_rho + se_rho, ymin = mean_rho - se_rho),
    size = 0.4, width = 0.3
  ) +
  theme_classic()

d_summarised %>% 
  ggplot(aes(x = bin, y = n)) +
  geom_bar(stat = 'identity') +
  theme_classic()
```

The overall trend looks pretty messy though:

```{r}
ggplot(d_bins, aes(x = log10(tract_size), y = rho_window)) +
  geom_point(alpha = 0.05, aes(color = bin)) +
  geom_smooth(method = 'lm') +
  coord_cartesian(y = c(0, 0.02))

```

KDEs for each log tract size with rho on the x and tract count on the y

```{r}
ggplot(d_bins, aes(x = rho_window, fill = bin, color = bin)) +
  geom_density(alpha = 0.2, position = 'stack') +
  coord_cartesian(x = c(0, 0.02))
```



Bin tract size on log scale and correlate with rho:

```{r}
d_bins_log <- d_bins %>% 
  mutate(log_size = log10(tract_size)) %>% 
  mutate(log_bin = floor(log_size)) %>% 
  group_by(log_bin) %>% 
  summarise(
    total_rho = sum(rho_vals),
    total_count = sum(rho_count),
    sd_rho = sd(rho_window),
    n = n()
  ) %>% 
  ungroup() %>% 
  mutate(
    mean_rho = total_rho / total_count,
    se_rho = sd_rho / sqrt(n)
  ) %>% 
  mutate(bin_start = as.integer(10 ^ log_bin),
         bin_end = lead(bin_start)) %>% 
  mutate(bin_end = ifelse(is.na(bin_end), Inf, bin_end)) %>% 
  mutate(bin_name = paste0(bin_start, '-', bin_end))

d_bins_log

d_bins_log %>% 
  filter(bin_start != 100000) %>% 
  ggplot(aes(x = bin_name, y = mean_rho)) +
  geom_point() +
  geom_errorbar(
    aes(ymin = mean_rho - se_rho, 
        ymax = mean_rho + se_rho)) +
  theme_bw() +
  labs(
    x = 'Tract size (bp)',
    y = expression(paste(rho))
  )

```

rewrite in terms of 'gene proximal' intergenic sequence

try with windowsizes of 1 kb and 2 kb

squelch bins >1000 kb into single bin (since so few tracts are > 10 kb)


## Rho in UTRs

```{r}
utr <- here('data/correlates/utr_tract_rho.tsv') %>% 
  read_tsv(col_types = cols()) %>% 
  filter(tract_size > 0)

utr3 <- here('data/correlates/utr3_tract_rho.tsv') %>% 
  read_tsv(col_types = cols())

utr_tracts <- left_join(utr, d_bins, 
  by = c('chrom', 'start', 'end', 'tract_size', 'rho_vals', 'rho_count', 'rho_window'))

utr3_full <- left_join(utr3, d_bins,
  by = c('chrom', 'start', 'end')) %>% 
  filter(!is.na(bin))
```

How does rho vary in UTRs by bin?

```{r}
# 5' UTRs
utr_tracts %>% 
  group_by(bin) %>% 
  summarise(
    total_rho = sum(utr_rho_vals),
    total_count = sum(utr_rho_count),
    sd_rho = sd(utr_rho_window),
    n = n()
  ) %>% 
  ungroup() %>% 
  mutate(
    mean_rho = total_rho / total_count,
    se_rho = sd_rho / sqrt(n)
  )

# how much is rho lower in UTRs next to tracts <4 kb than not?

utr_tracts %>% 
  mutate(bin = case_when(
    bin %in% c('short', 'medium') ~ '<4kb',
    bin %in% c('long', 'extended') ~ '>4kb'
  )) %>% 
  group_by(bin) %>% 
  summarise(
    total_rho = sum(utr_rho_vals),
    total_count = sum(utr_rho_count)
  ) %>% 
  ungroup() %>% 
  mutate(
    mean_rho = total_rho / total_count
  ) %>% 
  select(bin, mean_rho) %>% 
  spread(bin, mean_rho) %>% 
  mutate(percent_decrease = (`>4kb` - `<4kb`) / `>4kb`) # 27%

# 3' UTRs
utr3_full %>% 
  group_by(bin) %>% 
  summarise(
    total_rho = sum(utr3_rho_vals),
    total_count = sum(utr3_rho_count),
    sd_rho = sd(utr3_rho_window),
    n = n()
  ) %>% 
  ungroup() %>% 
  mutate(
    mean_rho = total_rho / total_count,
    se_rho = sd_rho / sqrt(n)
  )
  
```

In the 'extended' bin (ie no instances of 'both' sites) - mean rho at adjacent UTRs is well above the genomewide average!

Otherwise, rho is very low in most 5' UTRs, which matches the patterns seen in intergenic rho by tract length -
ie in shorter tracts, intergenic rho lower -> UTR rho lower, but in 'long' tracts, more upstream/downstream sites ->
higher rho -> higher rho in UTRs (trend most apparent in 'extended' bin)

similar trend in 3' UTRs, although long tract rho is less

```{r}
utr_bars <- utr_tracts %>% 
  group_by(bin) %>% 
  summarise(
    total_rho = sum(utr_rho_vals),
    total_count = sum(utr_rho_count),
    sd_rho = sd(utr_rho_window),
    n = n()
  ) %>% 
  ungroup() %>% 
  mutate(
    mean_rho = total_rho / total_count,
    se_rho = sd_rho / sqrt(n)
  )

utr3_bars <- utr3_full %>% 
  group_by(bin) %>% 
  summarise(
    total_rho = sum(utr3_rho_vals),
    total_count = sum(utr3_rho_count),
    sd_rho = sd(utr3_rho_window),
    n = n()
  ) %>% 
  ungroup() %>% 
  mutate(
    mean_rho = total_rho / total_count,
    se_rho = sd_rho / sqrt(n)
  )

utr_bars_all <- bind_rows(utr_bars, utr3_bars, .id = 'name') %>% 
  mutate(utr = case_when(
    name == 1 ~ 'five_prime',
    name == 2 ~ 'three_prime')
  )

ggplot(utr_bars_all, aes(x = bin, y = mean_rho, fill = utr)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  geom_errorbar(width = 0.4, position = position_dodge(width = 0.9),
    aes(ymin = mean_rho - se_rho, 
        ymax = mean_rho + se_rho)) +
  theme_minimal()


```


How does this look for percentile bins of tract length?

```{r}
utr5_tracts %>% 
  select(tract_size, utr_rho_window) %>%
  arrange(tract_size) %>% 
  mutate(tract_percentile = percent_rank(tract_size)) %>% 
  mutate(percentile_bin = floor(tract_percentile / 0.05) * 0.05) %>% 
  ggplot(aes(x = tract_percentile, y = utr_rho_window)) +
  geom_point() +
  geom_smooth()

```


## Hotspots in intergenic tracts

```{r}
intergenic_hotspots <- here('data/correlates/intergenic_tract_hotspots.tsv') %>% 
  read_tsv(col_types = cols()) %>% 
  filter(tract_size > 0) %>% 
  left_join(d_bins, by = c('chrom', 'start', 'end', 'rho_vals', 'rho_count', 'tract_size', 'rho_window'))
```

Hotspot counts/proportions by bin:

```{r}
intergenic_hotspots %>% 
  group_by(bin) %>% 
  summarise(
    avg_n_hotspots = mean(n_hotspots),
    avg_sites_in_hotspot = sum(sites_in_hotspot) / sum(rho_count),
    avg_upstream = mean(upstream_sites),
    avg_downstream = mean(downstream_sites),
    avg_both = mean(both_sites),
    avg_intergenic = mean(intergenic_sites)
  )
```

Hotspot strength:

```{r}
intergenic_hotspots %>% 
  filter(n_hotspots > 0) %>% 
  mutate(rate_ratio = rate_sum / n_hotspots) %>% 
  filter(rate_ratio < 2000) %>% # one freak hotspot
  group_by(bin) %>% 
  summarise(mean(rate_ratio))
```

does spatial distance between genes correlate with expression? 

## Varying flank sizes

Load in data:

```{r}
flank_100 <- read_tsv(here('data/correlates/intergenic_flanks_100.tsv'), col_types = cols()) %>% 
  left_join(select(d, -tract_size), by = c('chrom', 'start', 'end'))
flank_500 <- read_tsv(here('data/correlates/intergenic_flanks_500.tsv'), col_types = cols()) %>% 
  left_join(select(d, -tract_size), by = c('chrom', 'start', 'end'))
flank_1kb <- read_tsv(here('data/correlates/intergenic_flanks_1kb.tsv'), col_types = cols()) %>% 
  left_join(select(d, -tract_size), by = c('chrom', 'start', 'end'))
flank_2kb <- read_tsv(here('data/correlates/intergenic_flanks_2kb.tsv'), col_types = cols()) %>% 
  left_join(select(d, -tract_size), by = c('chrom', 'start', 'end'))
```


Split into bins of tracts <= 4 kb and > 4 kb:

```{r}
list(flank_500, flank_1kb, flank_2kb, flank_100) %>% 
  map_dfr(~ mutate(., type = case_when(
    tract_size <= 4000 ~ 'short',
    tract_size > 4000 ~ 'long'
  )) %>% 
  group_by(type) %>% 
  summarise(
    mean_left = mean(left_window),
    mean_right = mean(right_window)
  ), .id = 'name'
)

```

```{r}
flank_2kb %>% 
  filter(tract_size > 2000) %>% 
  select(-left_vals, -left_count, -left_window) %>% # remove downstream
  mutate(window = floor(tract_size / 2000) * 2000) %>% 
  mutate(window = ifelse(tract_size >= 10000, 10000, window)) %>% 
  group_by(window) %>% 
  summarise(total_right_vals = sum(right_vals), total_right_count = sum(right_count), n = n()) %>% 
  mutate(mean_right = total_right_vals / total_right_count) %>% 
  ggplot(aes(x = as.factor(window), y = mean_right)) +
  geom_point()
  
```


Seems the gene proximal regions are *way* higher in the longer tract flanks for all three windowsizes

Let's stick to the 2 kb window for simplicity

```{r}
flank_2kb %>% 
  mutate(type = case_when(
    tract_size <= 4000 ~ 'short',
    tract_size > 4000 ~ 'long'
  )) %>% 
  group_by(type) %>% 
  summarise(
    mean_left = mean(left_window),
    mean_right = mean(right_window)
  )
```

```{r}
# short
flank_2kb %>% 
  filter(tract_size <= 4000) %>% 
  select(contains('left'), contains('right')) %>% 
  summary()

# long
flank_2kb %>% 
  filter(tract_size > 4000) %>% 
  select(contains('left'), contains('right')) %>% 
  summary()
```

```{r}
flank_2kb %>% 
  mutate(type = case_when(
    tract_size <= 4000 ~ 'short',
    tract_size > 4000 ~ 'long'
  )) %>% 
  select(type, left_window, right_window) %>% 
  gather(window, value, -type) %>% 
  ggplot(aes(x = value, fill = type)) +
  geom_density(alpha = 0.2) +
  coord_cartesian(x = c(0, 0.25)) +
  facet_wrap(~ type)

# cdf
flank_2kb %>% 
  mutate(type = case_when(
    tract_size <= 4000 ~ 'short',
    tract_size > 4000 ~ 'long'
  )) %>% 
  select(type, left_window, right_window) %>% 
  gather(window, value, -type) %>% 
  ggplot(aes(x = value, color = type)) +
  stat_ecdf() +
  coord_cartesian(x = c(0, 0.1))
```

```{r}
# percent decrease
flank_2kb %>% 
  mutate(type = case_when(
    tract_size <= 4000 ~ 'short',
    tract_size > 4000 ~ 'long'
  )) %>% 
  group_by(type) %>% 
  summarise(
    total_rho_left = sum(left_vals),
    total_count_left = sum(left_count),
    total_rho_right = sum(right_vals),
    total_count_right = sum(right_count)
  ) %>% 
  ungroup() %>% 
  mutate(
    mean_rho_left = total_rho_left / total_count_left,
    mean_rho_right = total_rho_right / total_count_right
  ) %>% 
  select(type, mean_rho_left, mean_rho_right) %>% 
  gather(window, value, -type) %>% 
  spread(type, value) %>% 
  mutate(percent_decrease = (long - short) / long)

# log tract size plot
flank_log <- flank_2kb %>% 
  mutate(log_size = log10(tract_size)) %>% 
  mutate(log_bin = floor(log_size)) %>% 
  group_by(log_bin) %>% 
  summarise(
    total_rho_left = sum(left_vals),
    total_count_left = sum(left_count),
    sd_left = sd(left_window),
    total_rho_right = sum(right_vals),
    total_count_right = sum(right_count),
    sd_right = sd(right_window),
    n = n()
  ) %>% 
  ungroup() %>% 
  mutate(
    mean_rho_left = total_rho_left / total_count_left,
    mean_rho_right = total_rho_right / total_count_right,
    se_left = sd_left / sqrt(n),
    se_right = sd_right / sqrt(n)
  ) %>% 
  mutate(bin_start = as.integer(10 ^ log_bin),
         bin_end = lead(bin_start)) %>% 
  mutate(bin_end = ifelse(is.na(bin_end), Inf, bin_end)) %>% 
  mutate(bin_name = paste0(bin_start, '-', bin_end)) %>% 
  select(-bin_start, -bin_end, -starts_with('total_count'), -starts_with('total_rho')) %>% 
  gather(measure, value, -n, -bin_name, -log_bin) %>% 
  mutate(type = str_extract(measure, '(left|right)')) %>% 
  mutate(measure = str_replace(measure, '_(left|right)', '')) %>% 
  spread(measure, value)
  
flank_log %>%  
  ggplot(aes(x = bin_name, y = mean_rho, fill = type)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  geom_errorbar(aes(
    ymin = mean_rho - se,
    ymax = mean_rho + se),
    width = 0.4, position = position_dodge(width = 0.9)
  )

d_bins_log %>% 
  filter(bin_start != 100000) %>% 
  ggplot(aes(x = bin_name, y = mean_rho)) +
  geom_point() +
  geom_errorbar(
    aes(ymin = mean_rho - se_rho, 
        ymax = mean_rho + se_rho)) +
  theme_bw() +
  labs(
    x = 'Tract size (bp)',
    y = expression(paste(rho))
  )

```

Combine the two flanks into single data points:

```{r}
flank_log_combined <- flank_2kb %>% 
  mutate(log_size = log10(tract_size)) %>% 
  mutate(log_bin = floor(log_size)) %>% 
  mutate(rho_vals = left_vals + right_vals,
         rho_count = left_count + right_count) %>% 
  mutate(rho_window = rho_vals / rho_count) %>% 
  select(-left_vals, -right_vals, -left_count, 
         -right_count, -left_window, -right_window) %>% 
  group_by(log_bin) %>% 
  summarise(
    total_rho = sum(rho_vals),
    total_count = sum(rho_count),
    sd_rho = sd(rho_window),
    n = n()
  ) %>% 
  ungroup() %>% 
  mutate(
    mean_rho = total_rho / total_count,
    se_rho = sd_rho / sqrt(n)
  ) %>% 
  mutate(bin_start = as.integer(10 ^ log_bin),
         bin_end = lead(bin_start)) %>% 
  mutate(bin_end = ifelse(is.na(bin_end), Inf, bin_end)) %>% 
  mutate(bin_name = paste0(bin_start, '-', bin_end))

flank_log_combined %>% 
  filter(bin_start != 100000) %>% 
  ggplot(aes(x = bin_name, y = mean_rho)) +
  geom_point() +
  geom_errorbar(
    aes(ymin = mean_rho - se_rho, 
        ymax = mean_rho + se_rho)) +
  theme_bw() +
  labs(
    x = 'Tract size (bp)',
    y = expression(paste(rho))
  )
```


flank combined with overall rho by log tract size in same plot:

(for bins < 1000, 'flank' and 'tract' should be the same thing ofc)

```{r}
d_bins_log %>% 
  select(bin_name, bin_start, mean_rho, sd_rho, se_rho, n) %>% 
  mutate(type = 'tract') %>% 
  bind_rows(
    flank_log_combined %>% 
      select(bin_name, bin_start, mean_rho, sd_rho, se_rho, n) %>% 
      mutate(type = 'flank')
  ) %>% 
  select(type, everything()) %>% 
  filter(bin_start != 100000) %>% 
  ggplot(aes(x = bin_name, y = mean_rho, fill = type)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  geom_errorbar(
    aes(ymin = mean_rho - se_rho, 
        ymax = mean_rho + se_rho),
    width = 0.4, position = position_dodge(width = 0.9)) +
  theme_bw() +
  labs(
    x = 'Tract size (bp)',
    y = expression(paste(rho))
  ) +
  geom_hline(yintercept = 0.004086864, linetype = 'dashed') # genomewide mean
```

```{r}
flank_2kb %>% 
  filter(tract_size >= 10000, tract_size < 100000) %>% 
  filter(left_window < 0.1, right_window < 0.1) %>% 
  summary()
```

include non-flank rho as a third category:

```{r}
non_flank_2kb <- flank_2kb %>% 
  mutate(log_size = log10(tract_size)) %>% 
  mutate(log_bin = floor(log_size)) %>% 
  filter(tract_size > 4000) %>% 
  mutate(rho_vals = rho_vals - (left_vals + right_vals),
         rho_count = rho_count - (left_count + right_count))

non_flank_log_rho <- flank_2kb %>% 
  mutate(log_size = log10(tract_size)) %>% 
  mutate(log_bin = floor(log_size)) %>% 
  filter(tract_size > 4000) %>% 
  mutate(rho_vals = rho_vals - (left_vals + right_vals),
         rho_count = rho_count - (left_count + right_count)) %>% 
  select(chrom, start, end, tract_size, rho_vals, rho_count, log_size, log_bin) %>% 
  mutate(rho_window = rho_vals / rho_count) %>% 
  group_by(log_bin) %>% 
  summarise(
    total_rho = sum(rho_vals),
    total_count = sum(rho_count),
    sd_rho = sd(rho_window),
    n = n()
  ) %>% 
  ungroup() %>% 
  mutate(
    mean_rho = total_rho / total_count,
    se_rho = sd_rho / sqrt(n)
  ) %>% 
  mutate(bin_start = as.integer(10 ^ log_bin),
         bin_end = lead(bin_start)) %>% 
  mutate(bin_end = ifelse(is.na(bin_end), Inf, bin_end)) %>% 
  mutate(bin_name = paste0(bin_start, '-', bin_end)) %>% 
  mutate(type = 'non-flank')
```

Combine datasets:

```{r}
tracts_all <- d_bins_log %>% 
  select(bin_name, bin_start, mean_rho, sd_rho, se_rho, n) %>% 
  mutate(type = 'tract') %>% 
  bind_rows(
    flank_log_combined %>% 
      select(bin_name, bin_start, mean_rho, sd_rho, se_rho, n) %>% 
      mutate(type = 'flank')
  ) %>% 
  bind_rows(non_flank_log_rho) %>% 
  select(type, everything()) %>%
  filter(bin_start != 100000)

tracts_all
```

Getting 95% bootstrap intervals for each of these categories:

```{r}
boot_fxn <- function(d, rep_count, current_log_bin, data_type) {
  require(magrittr)
  boot_obj <- replicate(rep_count, {
    
    sampled_rows <- d %>% 
      mutate(log_size = log10(tract_size)) %>% 
      mutate(log_bin = floor(log_size)) %>% 
      mutate(bin_start = as.integer(10 ^ log_bin),
         bin_end = lead(bin_start)) %>% 
      mutate(bin_end = ifelse(is.na(bin_end), Inf, bin_end)) %>% 
      mutate(bin_name = paste0(bin_start, '-', bin_end)) %>% 
      filter(log_bin == current_log_bin) %>% 
      sample_frac(size = 0.05, replace = TRUE)
    
    out <- sum(sampled_rows$rho_vals) / sum(sampled_rows$rho_count)
    
    return(out)
  })
  
  message(paste('Completed sampling for bin', 10^current_log_bin))
  boot_obj <- na.omit(boot_obj)
  
  conf_df <- broom::tidy(t.test(boot_obj, alternative = 'two.sided', mu = 0, conf.level = 0.95)) %>% 
    select(estimate, p.value, conf.low, conf.high) %>% 
    mutate(bin_start = as.integer(10 ^ current_log_bin), type = data_type)
  return(conf_df)
}

boot_fxn(d, 1000, 2, 'tract')
```

Redoing for all categories:

```{r}
d_cis <- c(0:4) %>% 
  map_dfr(~ boot_fxn(d, 100, ., 'tract'))

flank_2kb_cis <- c(0:4) %>% 
  map_dfr(~ boot_fxn(
    flank_2kb %>% 
      mutate(rho_vals = rho_vals - (left_vals + right_vals),
      rho_count = rho_count - (left_count + right_count)),
    100, ., 'flank'
  ))

non_flank_2kb_cis <- c(3, 4) %>% 
  map_dfr(~ boot_fxn(non_flank_2kb, 100, ., 'non-flank'))
```

Writing CIs to files (since this takes a few minutes):

```{r}
write_csv(d_cis, here('data/correlates/d_cis.csv'))
write_csv(flank_2kb_cis, here('data/correlates/flank_2kb_cis.csv'))
write_csv(non_flank_2kb_cis, here('data/correlates/non_flank_2kb_cis.csv'))
```

Plotting:

```{r}
d_cis <- read_csv(here('data/correlates/d_cis.csv'), col_types = cols())
flank_2kb_cis <- read_csv(here('data/correlates/flank_2kb_cis.csv'), col_types = cols())
non_flank_2kb_cis <- read_csv(here('data/correlates/non_flank_2kb_cis.csv'), col_types = cols())

cis_all <- bind_rows(d_cis, flank_2kb_cis, non_flank_2kb_cis)

tracts_all_cis <- left_join(tracts_all, cis_all, by = c('bin_start', 'type'))
```


```{r}
tracts_all_cis %>% 
  ggplot(aes(x = bin_name, y = mean_rho, fill = type)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  geom_errorbar(
    aes(ymin = conf.low, 
        ymax = conf.high),
    width = 0.4, position = position_dodge(width = 0.9)) +
  theme_bw() +
  labs(
    x = 'Tract size (bp)',
    y = expression(paste(rho))
  ) +
  geom_hline(yintercept = 0.004086864, linetype = 'dashed') # genomewide mean
```














