---
title: "LD decay analysis"
author: Ahmed Hasan
output: pdf_document
---

Things to calculate:
1. Distance at which LD hits 0.5 * starting value
2. Distance at which LD decay 'flatlines' (dx becomes 0 to x s.d.)
3. (Plot lines)
4. Does the distance in (2) above correlate with chromosome length?

```{r}
require(knitr)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(fs)
```

Data preparation:

```{r}
fnames <- dir_ls('data/ld-decay/finals', regexp = 'chromosome\\_[0-9]{1,2}\\_fit\\.csv')
decay <- map_dfr(fnames, read_csv, col_types = cols())
```

## Distance at which LD = 0.5 * starting value

```{r}
get_halfdecay_point <- function(decay_df) {
  start_point <- decay_df %>% 
    head(n = 1) # get first row
  start_d <- start_point$d
  start_r2 <- start_point$r2
  chrom <- start_point$chrom
  
  half_r2 <- start_r2 * 0.5
  half_point <- decay_df %>% 
    arrange(d) %>% 
    filter(r2 >= half_r2) %>% 
    tail(n = 1) # get last row at r2 == half_r2
  
  end_d <- half_point$d
  end_r2 <- half_point$r2
  
  out <- tibble(
    chrom = chrom,
    start_d = start_d, end_d = end_d,
    start_r2 = start_r2, end_r2 = end_r2
  ) %>% 
    mutate(diff_d = end_d - start_d,
           diff_r2 = abs(end_r2 - start_r2))
  
  return(out)
}
```

Running this over the data:

```{r}
halfdecay <- decay %>% 
  split(.$chrom) %>% 
  map_dfr(~ get_halfdecay_point(.))

halfdecay
```

Stats:

```{r}
halfdecay %>% 
  select(diff_d) %>% 
  summary()
```



## Distance at which LD 'levels off'

```{r}
get_flatline_point <- function(decay_df) {
  start_point <- decay_df %>% head(n = 1)
  start_d <- start_point$d
  start_r2 <- start_point$r2
  chrom <- start_point$chrom
  
  flatline_point <- decay_df %>% 
    arrange(d) %>% 
    mutate(d2 = lead(d), r2_2 = lead(r2)) %>% 
    mutate(
      dx = abs(
        round((r2_2 - r2) / (d2 - d), digits = 5)
      )
    ) %>% 
    filter(dx > 0) %>% 
    tail(n = 1)
  
  end_d <- flatline_point$d
  end_r2 <- flatline_point$r2
  
  out <- tibble(
    chrom = chrom,
    start_d = start_d, end_d = end_d,
    start_r2 = start_r2, end_r2 = end_r2
  ) %>% 
    mutate(diff_d = end_d - start_d,
           diff_r2 = abs(end_r2 - start_r2))

  return(out)
}
```

Running over data:

```{r}
flatline_decay <- decay %>% 
  split(.$chrom) %>% 
  map_dfr(~ get_flatline_point(.))

flatline_decay
```

Stats:

```{r}
flatline_decay %>% 
  summarise(mean_r2 = mean(end_r2),
            sd_r2 = sd(end_r2),
            max_d = max(diff_d))
```


## Plot

Plot prep:

```{r}
library(wesanderson)
fig_colors <- as.character(wes_palette(17, name = 'GrandBudapest1', type = 'continuous'))

fig_1_theme <- theme(
  axis.title = element_text(family = "Helvetica", size = 16),
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(), 
  axis.text = element_text(family = "Helvetica", size = 16, color = 'black'),
  axis.line = element_line(colour = 'black', linetype = 'solid'),
  axis.line.x = element_line(size = 0.9),
  axis.line.y = element_line(size = 0.9),
  axis.ticks = element_line(colour = 'black', linetype = 'solid', size = 0.9),
  plot.tag = element_text(colour = 'black', size = 16, face = 'bold'),
  panel.background = element_blank()
)

lengths <- tibble(
  length = c(
    1922860, 3500558, 3826814, 4091191, 4157777, 5033832, 5206065, 6421821, 
    6576019, 7188315, 7783580, 7956127, 8033585, 9023763, 9219486, 9223677, 9730733),
  chrom = c(
     'chromosome_15', 'chromosome_5', 'chromosome_11', 'chromosome_4', 
     'chromosome_14', 'chromosome_8', 'chromosome_13', 'chromosome_7', 
     'chromosome_10', 'chromosome_17', 'chromosome_16', 'chromosome_9', 
     'chromosome_1', 'chromosome_6', 'chromosome_3', 'chromosome_2', 'chromosome_12')
)

```

Plot:

```{r}
decay %>% 
  left_join(lengths, by = 'chrom') %>% 
  mutate(d_kb = d / 1000) %>% 
  ggplot(aes(x = d_kb, y = r2, color = as.factor(chrom))) +
  fig_1_theme +
  geom_line() +
  coord_cartesian(x = c(0, 30)) +
  scale_color_manual(values = fig_colors) +
  guides(colour = FALSE) +
  labs(x = 'Distance between SNPs (kb)', 
       y = expression(paste('Linkage disequilibrium (r'^2, ')')))
```

## Decay and chromosome length

```{r}
flatline_decay %>% 
  left_join(lengths, by = 'chrom') %>% 
  ggplot(aes(x = length, y = diff_d)) +
  geom_point() +
  geom_smooth(method = 'lm')

flatline_decay %>% 
  left_join(lengths, by = 'chrom') %>% 
  cor.test(~ diff_d + length, data = ., method = 'spearman')
  
```


