---
title: "Hotspot analysis"
author: Ahmed Hasan
output: pdf_document
---

To do:
1. Do all chromosomes have hotspots? Which ones don't and which do?
2. How many hotspots are there? (+ summary stats listed below)

Summary stats:
- mean hotspot length
- mean rho fold increase over local background
- amount of sequence in hotspots (ie percent value of genome)
- mean distance b/w adjacent hotspots
- mean rho at hotspots

Will be using the block = 5 files for this (given simulation results)

```{r}
library(tidyverse)
library(fs)
library(here)
```

## Data prep

```{r}
non_overlapping <- function(d, windowsize) {
  d %<>%
    mutate(div = floor(block_start / windowsize), div2 = lead(div)) %>%
    filter(div == div2) %>%
    select(-contains('div'))
  return(d)
}

fnames <- dir_ls(here('data/ldhelmet/block_10/'), regexp = 'chromosome_[0-9]+_summarised\\.txt')
hotspots <- map(fnames, read_csv, col_types = cols()) %>% 
  map_dfr(~ filter(., rate_ratio >= 5))

hotspots_no_overlap <- hotspots %>% 
  non_overlapping(2000)

```


## Do all chromosomes have hotspots?

```{r}
unique(hotspots$chr) # yep
length(unique(hotspots$chr)) # yep
```

## Hotspot summary stats

```{r}
# time for a *shudder* for loop

assign_hotspot_groups <- function(d) {
  out <- numeric()
  counter <- 1
  d <- arrange(d, block_start)
  
  for (i in 1:(nrow(d) - 1)) {
    current_start <- d$block_start[i]
    next_start <- d$block_start[i + 1]
    if (next_start <= (current_start + 2000)) {
      out[i] <- counter
    } else if (next_start > (current_start + 2000)) {
      out[i] <- counter
      counter <- counter + 1
    }
  }
  
  # final case
  final_start <- d$block_start[nrow(d)]
  second_to_last_start <- d$block_start[nrow(d) - 1]
  if (final_start <= (second_to_last_start + 2000)) {
    out[nrow(d)] <- counter
  } else if (final_start > (second_to_last_start + 2000)) {
    out[nrow(d)] <- counter + 1
  }
  
  d_out <- d
  d_out$hotspot_group <- out
  return(d_out)
}
  
hotspot_groups <- hotspots %>% 
  split(.$chr) %>%
  map(~ assign_hotspot_groups(.))

```


Summary stats:
- hotspot counts
- mean hotspot length
- amount of sequence in hotspots (ie percent value of genome)
- mean rho at hotspots + fold increase over local background
- mean distance b/w adjacent hotspots

### Hotspot counts

```{r}
hotspot_counts <- hotspot_groups %>% 
  map_dfr(~ summarise(., n = max(hotspot_group)), .id = 'name')

sum(hotspot_counts$n) # 875
```

### Mean hotspot length

```{r}
hotspot_groups %>% 
  map_dfr(
    ~ group_by(., hotspot_group) %>% 
      summarise(start = min(block_start), end = max(block_end)) %>% 
      mutate(length = end - start),
    .id = 'name'
  ) %>% 
  summarise(mean_length = mean(length)) # 3428 bp
```

### Amount of sequence in hotspots

```{r}
# total amt of sequence
genome_length <- map(fnames, read_csv, col_types = cols()) %>% 
  map_dfr(~ summarise(., length = max(block_end)), .id = 'chr') %>% 
  select(length) %>% 
  summarise(total = sum(length)) %>% 
  unlist()

genome_length
```

```{r}
# amt of sequence just in hotspots
hotspot_sequence <- hotspot_groups %>% 
  map_dfr(
    ~ group_by(., hotspot_group) %>% 
      summarise(start = min(block_start), end = max(block_end)) %>% 
      mutate(length = end - start),
    .id = 'name'
  ) %>% 
  summarise(total = sum(length)) %>% 
  unlist()

hotspot_sequence / genome_length # 0.0275
```

### Mean rho + fold increase over local background

```{r}
hotspot_groups %>% 
  map_dfr(
    ~ group_by(., hotspot_group) %>% 
      summarise(
        mean_hotspot_rho = mean(block_rate),
        hotspot_fold_increase = mean(rate_ratio)
      ),
    .id = 'chr'
  ) %>% 
  summarise(
    mean_rho = mean(mean_hotspot_rho),
    fold_increase = mean(hotspot_fold_increase)
  )
```

### Mean distance between adjacent hotspots

```{r}
hotspot_distances <- hotspot_groups %>% 
  map_dfr(
    ~ group_by(., hotspot_group) %>% 
      summarise(
        start = min(block_start),
        end = max(block_end)
      ) %>% 
      mutate(next_start = lead(start)) %>% 
      mutate(distance = next_start - end),
    .id = 'chr'
  )

hotspot_distances %>% 
  summarise(mean_distance = mean(distance, na.rm = TRUE)) # 122.05 kb
```

