---
title: "Recombination landscape analysis"
author: Ahmed Hasan
output: pdf_document
---

To do:
1. What is the extent of recombination variation across chromosomes?
2. Do shorter chromosomes have higher recombination rates?

Will be using the block = 100 files for this -- block = 5 is for hotspot analysis

```{r setup, include=FALSE}
require(knitr)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```


```{r}
library(tidyverse)
library(fs)
```

## Extent of rho variation across chromosomes

```{r}
non_overlapping <- function(d, windowsize) {
  d %<>%
    mutate(div = floor(block_start / windowsize), div2 = lead(div)) %>%
    filter(div == div2) %>%
    select(-contains('div'))
  return(d)
}

fnames <- dir_ls('data/ldhelmet/block_100/', regexp = 'chromosome_[0-9]+_summarised\\.txt')
landscape <- map(fnames, read_csv, col_types = cols()) %>% 
  map_dfr(~ non_overlapping(., 2000))
```

Summary stats:

```{r}
landscape %>% 
  group_by(chr) %>% 
  summarise(mean_rho = mean(block_rate, na.rm = TRUE)) %>% 
  arrange(mean_rho)
```

## Rho and chromosome lengths

```{r}
lengths <- tibble(
  length = c(
    1922860, 3500558, 3826814, 4091191, 4157777, 5033832, 5206065, 
    6421821, 6576019, 7188315, 7783580, 7956127, 8033585, 9023763, 
    9219486, 9223677, 9730733),
  chr = c(
     'chromosome_15', 'chromosome_5', 'chromosome_11', 'chromosome_4', 
     'chromosome_14', 'chromosome_8', 'chromosome_13', 'chromosome_7', 
     'chromosome_10', 'chromosome_17', 'chromosome_16', 'chromosome_9', 
     'chromosome_1', 'chromosome_6', 'chromosome_3', 'chromosome_2', 'chromosome_12')
)

rho_lengths <- landscape %>% 
  group_by(chr) %>% 
  summarise(mean_rho = mean(block_rate, na.rm = TRUE)) %>% 
  arrange(mean_rho) %>% 
  left_join(lengths, by = 'chr')
```

Plot:

```{r}
rho_lengths %>% 
  ggplot(aes(x = length, y = mean_rho)) +
  geom_point() +
  geom_smooth(method = 'lm')
```

Linear fit:

```{r}
rho_lengths %>% 
  lm(mean_rho ~ length, data = .) %>% 
  summary()
```

Comparison with decay R:

```{r}
fitted_rho <- read_csv('data/ld-decay/finals/fitted_rho.csv', col_types = cols()) %>% 
  rename(chr = chrom) %>% 
  select(-term)

joined <- rho_lengths %>% 
  left_join(fitted_rho, by = 'chr')

joined %>% 
  ggplot(aes(x = mean_rho, y = estimate)) +
  geom_point() +
  geom_smooth(method = 'lm')
```

chromosome 15's dramatically high mean rho makes it quite the outlier...


```{r}
# without chromosome 15?
joined %>% 
  filter(chr != 'chromosome_15') %>% 
  ggplot(aes(x = mean_rho, y = estimate)) +
  geom_point() +
  geom_smooth(method = 'lm', se = FALSE) +
  geom_abline(slope = 1, intercept = 0) +
  coord_cartesian(x = c(0.002, 0.01), y = c(0.002, 0.01))

```

```{r}
joined %>% 
  filter(chr != 'chromosome_15') %>% 
  lm(estimate ~ mean_rho, data = .) %>% 
  summary()
```


